Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-17 E-Hypergraphs-LICS24.tex
EDPOI rewrite defn

CHRIS: FIX BELOW DEFINITION. 
\begin{definition}[Convex EDPOI rewriting]
\label{def:dpoi-e}
Given an extended span of morphisms 
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
\]
in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule} 
\[
    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] 
if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram such that the two marked squares are pushouts.
\[
 \scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_square}
 }
\]
 and the following conditions hold:
    \begin{enumerate}
        \label{dpoi-e:assumptions}
        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
        \item $m(L)$ is a down-closed e-hypergraph;
        \item all elements in the image of $i+j$ share the same set of parents;
        \item $m(v)  \consistency m(w)$ for all $v,w$ in the image of $i+j$.
        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
	\item the internal interfaces are given by the following:
		\begin{align*}
		&n'' = (i'' \setminus i) + l\\
		&k'' = (j'' \setminus j) + r\\
		&n'' \to \mathcal{H} = [f_1,f_2] \text{ where }\\
		&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H}\\ 
		&f_2 : l \to \mathcal{C} \to \mathcal{H} 
		\end{align*}
    \end{enumerate}
% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
%  \[
%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
% \]
Since both $f_1$ and $f_2$ are monos, their copairing is also a mono. 
Similarly for $k'' \to \mathcal{H}$.

>>>>

CHRIS: FIX BELOW DEFINITION. 
\begin{definition}[Convex EDPOI rewriting]
\label{def:dpoi-e}
Given an extended span of morphisms 
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k \xrightarrow{} n'' + k'' \xrightarrow{} \mathcal{H}
\]
in $\catname{EHyp(\Sigma)}$, we say $\mathcal{G}$ \textit{rewrites (convexly) to} $\mathcal{H}$ (\textit{with external interface} $n + k$ and \textit{taking internal interface} $n'+k'$ \textit{to} $n'' + k''$) --- denoted by $\mathcal{G} \Rrightarrow \mathcal H$  --- \textit{via a rewrite rule} 
\[
    \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] 
if there exists an object $\mathcal{C}$ and morphisms which complete the following commutative diagram 
\[
 \scalebox{0.75}{
    \tikzfig{combinatorial_semantics/DPOI_square}
 }
\]
such that the two marked squares are pushouts and the following conditions hold:
    \begin{enumerate}
        \label{dpoi-e:assumptions}
        \item $m : \mathcal{L} \to \mathcal{G}$ is a convex match;
        \item $m(L)$ is a down-closed e-hypergraph;
        \item all elements in the image of $i+j$ share the same set of parents;
        \item $m(v)  \consistency m(w)$ for all $v,w$ in the image of $i+j$.
        \item $i + j \to \mathcal{C} \to \mathcal{G}$ is a boundary complement;
	\item the internal interfaces are given by the following:
		\begin{align*}
		&n'' = (i'' \setminus i) + l\\
		&k'' = (j'' \setminus j) + r\\
		&n'' \to \mathcal{H} = [f_1,f_2] \text{ where }\\
		&f_1 : i'' \setminus i \to \mathcal{R} \to \mathcal{H}\\ 
		&f_2 : l \to \mathcal{C} \to \mathcal{H} 
		\end{align*}
    \end{enumerate}
% The internal interfaces of the resulting e-hypergraph $\mathcal{H}$ from the cospan
%  \[
%     \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k
% \]
Since both $f_1$ and $f_2$ are monos, their copairing is also a mono. 
Similarly for $k'' \to \mathcal{H}$.
